This is Project Assignment 1 for course "Reproducable Research"
==============================================================


```{r setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```


```{r Loading the data}
setwd("C:/Users/ESRAAD/Desktop/Reproducable Research/Week2/")
data <- read.csv(paste0("repdata_data_activity/activity.csv"))
```


```{r Transform the data}
data$date <- as.Date(as.character(data$date), origin = "1899-12-30")
```

```{r Question 1 - Mean Total Number of Steps Taken per day}
library(stats)
Total_Steps <- aggregate(steps ~ date, data, FUN = "sum")
hist(Total_Steps$steps, xlab = "Steps taken per day", ylab = "Number of Days", main = "Total Number of seps taken by day")
mean <- mean(Total_Steps$steps)
median <- median(Total_Steps$steps)
```
- Mean of the number of steps taken per day is `r mean`.
- Median of the number of steps taken per day is `r median`.


```{r Question 2 - Average daily activity pattern}
library(ggplot2)
Average_Steps <- aggregate(steps ~ interval, data, FUN = "mean")
with(Average_Steps, plot(interval, steps, type = "l", xlab = "Daily Intervals", ylab = "Average Number of Steps"))
```


```{r Question 3 - Imputing Missing Values}
data1 <- data
Missing <- is.na(data1$steps)
NumberMissing <- sum(Missing)
m <- mean(data1$steps, na.rm = T)
data1$steps[Missing] <- m
head(data1)
Total_Steps1 <- aggregate(steps ~ date, data1, FUN = "sum")
hist(Total_Steps1$steps, xlab = "Steps taken per day", ylab = "Number of Days", main = "Total Number of seps taken by day")
mean1 <- mean(Total_Steps1$steps)
median1 <- median(Total_Steps1$steps)
```
- Number of missing values is `r NumberMissing` which will be replaced by the mean which is `r m`.
- Mean of the number of steps taken per day after imputing missing values is `r mean1`.
- Median of the number of steps taken per day after imputing missing values is `r median1`.

```{r Question 4 - Differences in activity patterns between weekdays and weekends}
library(dplyr)
library(lattice)
data2 <- data1
data2$DayType <- factor(c("Weekday", "Weekend"))
Weekend <- c("Sunday", "Saturday")
data2$DayType <- ifelse(weekdays(data2$date) %in% Weekend, "Weekend", "Weekday")
Steps_DayType <- group_by(data2, interval, DayType)
Average_Steps_DayType <- summarize(Steps_DayType, mean = mean(steps))
xyplot(mean ~ interval | DayType, data = Average_Steps_DayType, layout = c(1,2), type = "l", xlab = "Daily Intervals", ylab = "Average Number of steps")
```



